Waving and swaying

### 2014-08-28 Paper/grass.tex
figure label

The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point in the flow profile. We find that this inflection point
is located near the tip of grass canopy. Our analysis shows that flow in presence of grass become unstable not only through a mechanism of Kelvin-Helmholtz instability associated with 
inflection point but also through shear instability of flow above grass coupled with the porous flow through grass bed.
\end{abstract}
\maketitle

@@

% Our calculation of linear stability analysis of flow in presence of grass shows that flow through vegetation become unstable not only through a mechanism of Kelvin-Helmholtz instability associated with strong shear near the canopy top but also through an instability of the unvegetated flow coupled with flow deep within the vegetation. We call this new instability the unvegetated layer instability for brevity. We further show that inclusion of drag due to grass predicts a threshold condition for waving. We confirm our analysis by comparing prediction from theory with field and experimental observations.

@@

continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} \shreyas{Where did this model come from? Has such a model been used before? Citations to those needed.} as

@@

$P$ the pressure, $\mu$ the turbulent viscosity and $\mathbf{g}$ the acceleration due to gravity. The Reynolds number of the flow on the scale of a grass blade is expected to be O($10^2$-$10^3$), and hence the drag force is dominated by form drag, and is modeled 

@@

normal to grass while $\hat{t},\hat{n}$ being unit vector along and normal to grass. We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. In the field, both $C_N \& N_g$ vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N$ \& $N_g$ to be constants.

@@

\caption{Basic flow profile (left )from the experiment and model and the model of canopy (bottom) along with variation of shear layer (inset) }

@@

Momentum and mass balance equation can be non-dimensionlize using half channel height $H$, associated advection time H/$U_0$ and velocity $U_0$. We are left with three non-dimensional parameters in the model, viz. the Reynolds number $R= {\rho U_0 H}/{\mu}$, the drag parameter $\delta/H$, and the submergence ratio of the grass $h_g/H$. With these scaling along with the use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance; 

@@

Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$,  we get modified Orr-Sommerfield equation. The eigenvalue $\sigma$ for a given $k$ provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$.

>>>>

The spontaneous waving of marine grass is thought to be due to a Kelvin-Helmholtz instability resulting from an inflection point in the flow profile. We find that this inflection point
is located near the tip of grass canopy. Our analysis shows that flow in presence of grass become unstable not only through a mechanism of Kelvin-Helmholtz instability associated with 
inflection point but also through shear instability of flow above grass coupled with the porous flow through grass bed. We further show that in canopy with high submergence ratio
Kelvin-Helmholtz insatbility is damped due to grass drag.

\end{abstract}
\maketitle

@@

% Our calculation of linear stability analysis of flow in presence of grass shows that flow through vegetation become unstable not only
%through a mechanism of Kelvin-Helmholtz instability associated with strong shear near the canopy top but also through an instability of the 
%unvegetated flow coupled with flow deep within the vegetation. We call this new instability the unvegetated layer instability for brevity. We further show that inclusion of drag due to grass predicts a threshold condition for waving. We confirm our analysis by comparing prediction from theory with field and experimental observations.

@@

continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance\cite{Vivoni98,Nepf99,Ghisal02,Delangre04,Delangre06} 
\shreyas{Where did this model come from? Has such a model been used before? Citations to those needed.} as

@@

$P$ the pressure, $\mu$ the turbulent viscosity and $\mathbf{g}$ the acceleration due to gravity. The Reynolds number of the flow on the scale of a grass blade is expected to 
be O($10^2$-$10^3$), and hence the drag force is dominated by form drag, and is modeled 

@@

normal to grass while $\hat{t},\hat{n}$ being unit vector along and normal to grass. We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. In the field, both $C_N \& N_g$ 
vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N$ \& $N_g$ to be constants.

@@

\caption{Base flow profile (left )from the experiment of Vivoni\cite{Vivoni98} (Case-3A (Figure-3.6A) with 330 plants/$m^2$, 6 blades per plant, 
plant height = 16.0$\pm0.2$ and blade width of 0.28cm)
 and it's approximation with $U_0=31cm/s$ and $\delta = 0.2H$ with our model. The model of canopy is shown in the bottom and the variation of shear layer in the inset }

@@

Momentum and mass balance equation can be non-dimensionlize using half channel height $H$, associated advection time H/$U_0$ and velocity $U_0$. 
We are left with three non-dimensional parameters in the model, viz. the Reynolds number $R= {\rho U_0 H}/{\mu}$, the drag parameter $\delta/H$, 
and the submergence ratio of the grass $h_g/H$. With these scaling along with the use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$ to satisfy mass balance; 

@@

Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$,  
we get modified Orr-Sommerfield equation. The eigenvalue $\sigma$ for a given $k$ provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$.
